Alien species

### 2023-03-21 DRAFT_alien_species.Rmd
ea_spread

@@

In addition to looking at the area weighted means and its relationship
with the pressure indicator (HIF), we can do the same at the polygon
level. Because the indicator is pseudo-continuous, boxplots for example
become useless, so I will present a relative frequency plot.

@@

<!-- FORTSETT HER -->

I want to add a threshold so that we don't end up over extrapolating
based on too few data points. I have previousl used 150 data pionts, and I think that was about right.

```{r, eval=F}
wetlands_slitasje_extr <- ea_spread(indicator_data = wetlands,
                         indicator = i_2020,
                         regions = HIA_regions,
                         groups = region_HIF,
                         threshold = 150)

seminat_slitasje_extr <- ea_spread(indicator_data = seminat,
                         indicator = i_2020,
                         regions = HIA_regions,
                         groups = region_HIF,
                         threshold = 150)

natOpen_slitasje_extr <- ea_spread(indicator_data = natOpen,
                         indicator = i_2020,
                         regions = HIA_regions,
                         groups = region_HIF,
                         threshold = 150)
#saveRDS(wetlands_slitasje_extr, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/wetlands_slitasje_extr.rds")
#saveRDS(seminat_slitasje_extr,  "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/seminat_slitasje_extr.rds")
#saveRDS(natOpen_slitasje_extr,  "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/natOpen_slitasje_extr.rds")
```

```{r, include=F}
wetlands_slitasje_extr <- readRDS(paste0(pData, "cache/wetlands_slitasje_extr.rds"))
seminat_slitasje_extr  <- readRDS(paste0(pData, "cache/seminat_slitasje_extr.rds"))
natOpen_slitasje_extr  <- readRDS(paste0(pData, "cache/natOpen_slitasje_extr.rds"))
```

It's easier to see what's happening if we zoom in a bit. Lets get a
boundary box around Trondheim.

```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(wetlands_slitasje_extr))
```

Cropping the data to the bbox

```{r, warning=F}
wetlands_slitasje_extr_trd <- sf::st_crop(wetlands_slitasje_extr, myBB)
seminat_slitasje_extr_trd <- sf::st_crop(seminat_slitasje_extr, myBB)
natOpen_slitasje_extr_trd <- sf::st_crop(natOpen_slitasje_extr, myBB)
```

```{r extrapolated-slitasje, fig.cap="Slitasje indicator extrapolated over Trondheim"}
myCol <- "RdYlGn"
tmap_arrange(
tm_shape(wetlands_slitasje_extr_trd)+
  tm_polygons(col = "w_mean",
    title="Slitasje indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="blue")+
  tm_layout(title = "Wetland"),

tm_shape(seminat_slitasje_extr_trd)+
  tm_polygons(col = "w_mean",
    title="Slitasje indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="blue")+
  tm_layout(title = "Semi-natural areas"),

tm_shape(natOpen_slitasje_extr_trd)+
  tm_polygons(col = "w_mean",
    title="Slitasje indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="blue")+
  tm_layout(title = "Naturally open areas"),
HIA_trd)
```

Note that for wetland and semi-natural areas we did not
extrapolate values for the most urban areas, and that for naturally open
areas we did not extrapolate to the least affected areas.
For the other regions we generally excluded more HIFs.

> *The maps above are easy to misunderstand, and they should therefor be
> communicated with caution. They should be considered as temporary
> files or as by-products.*

>>>>

Here is a similar figure, looking at the relative frequency of indicator level of polygons and within each HIA.

@@

@@

I want to add a threshold so that we don't end up over extrapolating
based on too few data points. I will use 20 data points as a minimum.

```{r, eval=F}
wetland_alien_extr <- ea_spread(indicator_data = wetlands,
                         indicator = i_2020,
                         regions = HIA_muni,
                         groups = muni_HIF,
                         threshold = 20)

seminat_alien_extr <- ea_spread(indicator_data = seminat,
                         indicator = i_2020,
                         regions = HIA_muni,
                         groups = muni_HIF,
                         threshold = 20)

natOpen_alien_extr <- ea_spread(indicator_data = natOpen,
                         indicator = i_2020,
                         regions = HIA_muni,
                         groups = muni_HIF,
                         threshold = 20)
#saveRDS(wetlands_alien_extr, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/wetlands_alien_extr.rds")
#saveRDS( seminat_alien_extr,  "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/seminat_alien_extr.rds")
#saveRDS( natOpen_alien_extr,  "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/natOpen_alien_extr.rds")

#saveRDS(wetlands_alien_extr,  "P:/41201785_okologisk_tilstand_2022_2023/data/cache/wetlands_alien_extr.rds")
#saveRDS( seminat_alien_extr,  "P:/41201785_okologisk_tilstand_2022_2023/data/cache/seminat_alien_extr.rds")
#saveRDS( natOpen_alien_extr,  "P:/41201785_okologisk_tilstand_2022_2023/data/cache/natOpen_alien_extr.rds")

```{r, include=F}
wetland_alien_extr <- readRDS(paste0(pData, "cache/wetlands_alien_extr.rds"))
seminat_alien_extr  <- readRDS(paste0(pData,  "cache/seminat_alien_extr.rds"))
natOpen_alien_extr  <- readRDS(paste0(pData,  "cache/natOpen_alien_extr.rds"))
```

It's easier to see what's happening if we zoom in a bit.

Trondheim is municipality nr 5001

3020 is Nordre Follo

5418 is Målselv

```{r}
wetland_alien_extr <- wetland_alien_extr %>%
  separate(ID,
           into = c("ID", "municipalityNumber", "HIF"),
           sep = "_") %>%
  add_column(eco = "wetlands")

seminat_alien_extr <- seminat_alien_extr %>%
  separate(ID,
           into = c("ID", "municipalityNumber", "HIF"),
           sep = "_") %>%
  add_column(eco = "semi_natural")

natOpen_alien_extr <- natOpen_alien_extr %>%
  separate(ID,
           into = c("ID", "municipalityNumber", "HIF"),
           sep = "_") %>%
  add_column(eco = "naturally_open")

alien_extr <- rbind(
  wetland_alien_extr,
  seminat_alien_extr,
  natOpen_alien_extr
)
```

```{r}
alien_extr_trd_nf <- alien_extr %>%
  filter(municipalityNumber %in% c("5001", "3020", "5418"))
```

```{r extrapolated-alien, fig.cap="Alein species indicator extrapolated over Nordre Follo (3020), Trondheim (5001) and Målselv (5418). Note that the maps are not masked by ecosystem type, so the three ecosystem specific indicators are overlaping. The bottom row shows the location of the four different human impact sones."}
myCol <- "RdYlGn"

tmap_arrange(
tm_shape(alien_extr_trd_nf)+
  tm_polygons(col = "w_mean",
    title="Alien species indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_facets(by=c("eco", "municipalityNumber"))
,
tm_shape(alien_extr_trd_nf)+
  tm_polygons(col = "HIF",
    title="Human impact factor",
    palette = "-viridis",
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_facets(by=c("municipalityNumber"))
,
ncol=1,
heights = c(1,0.3))
```

Figure \@ref(fig:extrapolated-alien) shows that this indicator will have a lot of missing data. For Nordre Follo and Trondheim for example, we do not have any indicator values tied to naturally open or wetland ecosystems, and for semi natural ecosystems we only have data for HIF class 2. This is, however, still a lot better than discarding the entire data set due to issues related to area representativity. The amount of missing data in the indicator maps is closely ties to the chosen threshold value for how many data points we must have in order to calculate an (area weighted) men indicator value. We have used 20 as the threshold here. For Målselv (far right column), we ended up having enough data for wetlands in HIF class 0, and since most of the municipality is in this HIF class we get a lot of indicator coverage.
